TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-08-17 paper/draft.tex
note
table header description

Another open question is about the [Fe/H] validation, and whether we are happy with just the cmd plot of metallicity. I think our ideas for other tests were:
\begin{itemize}
\item Looking at M+M binary pairs: there are 174 pairs and it looks by eye that most of the pair lines are parallel to the main sequence. I don't know if there's some more clever way to see if the pairs lie in the same metallicity space of the cmd, so this is probably not that useful
\item Overplotting some isochrones at varying metallicities--which I'm still working on
\end{itemize}
}\color{black}

Also I still have to make the vertical action plot.
\end{abstract}

@@

\begin{longrotatetable}
\begin{deluxetable*}{cccccccccc}
\tabletypesize{\scriptsize}
\tablecaption{Temperatures/Metallicities (from Mann-trained model), and spectral types (from West-trained model) for the sample 10,311 APOGEE M dwarfs, as well as goodness of fit for the respective models. \label{test_results_table}}
\tablehead{\colhead{APOGEE ID} & \colhead{Gaia\ ID} & \colhead{RA}  & \colhead{DEC} & \colhead{\teff} & \colhead{[Fe/H]}  & \colhead{SPT} & \colhead{$\chi^2$ Mann} & \colhead{$\chi^2$ West} & \colhead{\nodata} \\
 &  & (deg)  & (deg) & (K) & (dex) &  &  &  & }   
\startdata
2M00432365+0332534 & 413814124681430528  & 10.848967  & 3.548273   & 3417 & 0.18  & 3.0 & 7910  & 7912   & \nodata \\
2M10382316+2457051 & 808931427773946112  & 159.596215 & 24.951373  & 3358 & -0.25 & 3.4 & 19051 & 25518  & \nodata \\
2M09364698+5602332 & 598955557617293440  & 144.195513 & 56.042461  & 3819 & 0.01  & 1.5 & 7981  & 6894   & \nodata \\
2M12165176+1543232 & 1538883006881701120 & 184.215466 & 15.723067  & 3948 & 0.16  & 0.7 & 25354 & 19983  & \nodata \\
2M18501195+4809327 & 2037435595069990016 & 282.549868 & 48.15976   & 3899 & 0.24  & 0.9 & 7205  & 7151   & \nodata \\
2M07572656+4703318 & 682241540493118080  & 119.360502 & 47.05852   & 3636 & -0.17 & 2.4 & 38430 & 45625  & \nodata \\
2M11083992+1112503 & 1536026200794333696 & 167.166344 & 11.213979  & 3949 & 0.1   & 0.8 & 21165 & 17506  & \nodata \\
2M12135300+1419452 & 3695183295027782656 & 183.470871 & 14.32924   & 3824 & 0.13  & 1.3 & 35018 & 21010  & \nodata \\
2M12273816-0049437 & 3682549562826659456 & 186.909269 & -0.828949  & 3715 & 0.04  & 1.8 & 14179 & 10656  & \nodata \\
2M10513639-0150064 & 777824835073481984  & 162.901465 & -1.835273  & 3997 & -0.15 & 1.2 & 80915 & 75157  & \nodata \\
2M02442390-1352593 & 447040266556331008  & 41.099824  & -13.883403 & 3676 & -0.28 & 2.3 & 41375 & 56704  & \nodata \\
2M06281785+0940439 & 65792100044278400   & 97.074558  & 9.678234   & 3659 & -0.29 & 2.3 & 78827 & 118883 & \nodata \\
2M13441275+5538012 & 1551387099791031552 & 206.053693 & 55.633288  & 3856 & -0.1  & 1.6 & 50070 & 45895  & \nodata \\
2M12473762+1107449 & 1251277737449848064 & 191.906593 & 11.129187  & 3244 & -0.42 & 4.0 & 30184 & 47020  & \nodata \\
2M01274162+8435576 & 523638809262601216  & 21.924655  & 84.599233  & 3534 & 0.07  & 2.6 & 15529 & 15113  & \nodata \\
2M13464494+4034158 & 1247733770955275008 & 206.686943 & 40.571056  & 3731 & -0.16 & 2.1 & 53196 & 50765  & \nodata \\
2M00504161+0219408 & 2376325179183866240 & 12.673481  & 2.327846   & 3745 & 0.4   & 1.4 & 76966 & 52035  & \nodata \\
2M14494090+3201268 & 1292097793826595328 & 222.420454 & 32.023994  & 3228 & 0.0   & 3.9 & 28733 & 38112  & \nodata \\
2M16434969-0123117 & 1406510018874952960 & 250.957083 & -1.386602  & 3669 & 0.03  & 2.0 & 22346 & 17186  & \nodata \\
2M13444356+1655384 & 1565587017583214464 & 206.18144  & 16.927287  & 3652 & 0.13  & 2.1 & 47828 & 40933  & \nodata \\
2M08162613+3250515 & 3053990410555781120 & 124.108745 & 32.84736   & 3837 & -0.15 & 1.8 & 30389 & 35103  & \nodata \\
2M10414608+2435412 & 3758656623830055168 & 160.441928 & 24.594731  & 4020 & -0.19 & 1.1 & 41857 & 36934  & \nodata \\
2M14373854+5014502 & 1188221157913600384 & 219.410555 & 50.247523  & 3380 & -0.13 & 3.2 & 35397 & 41989  & \nodata \\
\nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata
\enddata
\end{deluxetable*}
\end{longrotatetable}
\begin{deluxetable*}{cccccccccc}
\tablecaption{Positions and 3D kinematics for 10,311 M dwarfs; parallaxes and proper motions from \gaia\ + radial velocities from \apogee.}
\tablehead{\colhead{\nodata} & \colhead{$\varpi$} & \colhead{$\sigma_\varpi$} & \colhead{$\mu_\alpha$}  & \colhead{$\sigma_{\mu\alpha}$} & \colhead{$\mu_\delta$} & \colhead{$\sigma_{\mu\delta}$}  & \colhead{$v_{los}$} & \colhead{\nodata} \\
 & (mas) & (mas) & (mas/yr)  & (mas/yr) & (mas/yr) & (mas/yr)  & (km/s) & \colhead{\nodata}}   
\nodata & 0.4095971279793963  & 0.018344701025907368 & 94.6092250382206    & 0.2326543993399578  & 21.139312054788142  & 0.10645732514148187  & -14.8169    & \nodata \\
\nodata & 0.7907713253583666  & 0.05017993250995411  & -61.145608504420785 & 0.19267235934981508 & -9.292283439334902  & 0.191605833991544    & 30.0646     & \nodata \\
\nodata & 2.437411282195209   & 0.06484152349097437  & -40.89370712477501  & 0.12447374577648344 & -27.01010381099313  & 0.12376849807276755  & -25.7591    & \nodata \\
\nodata & 4.748695834113324   & 0.025884382476497582 & -39.804579591840216 & 0.0930132037970028  & -10.36148239898068  & 0.057746591921984974 & -7.28313    & \nodata \\
\nodata & 0.3830610102451543  & 0.020679199019227986 & 8.768467380162365   & 0.03404660389361428 & 142.96097527921128  & 0.03222662578521854  & -67.5318    & \nodata \\
\nodata & 1.2156916470683867  & 0.05850523666981387  & -34.02970707374436  & 0.5249923050897877  & -60.82222958247146  & 0.41961695655028974  & 27.5802     & \nodata \\
\nodata & 3.2562530226439863  & 0.12635148166126747  & -14.18737757924032  & 0.08323990329349258 & 3.839844523005405   & 0.08709819792316187  & 1.19813     & \nodata \\
\nodata & 1.1214563311692185  & 0.029138557090469957 & 3.8612047002174377  & 0.10975162725538208 & -6.6212605579145185 & 0.0625749242163525   & -12.022     & \nodata \\
\nodata & 0.4337056924769619  & 0.03519634806417813  & 63.58287541662625   & 0.12216317236007974 & -30.844101370573288 & 0.06988132634903554  & 5.3206      & \nodata \\
\nodata & 4.860871276791149   & 0.05422112071607727  & -38.48480428052528  & 0.06681092801786047 & -46.03340135581453  & 0.05013526442486122  & 5.74919     & \nodata \\
\nodata & 0.3900296037356137  & 0.022457415816365067 & 48.72202372576884   & 0.09931020159590297 & -41.61843538517975  & 0.0939261318266816   & 56.5766     & \nodata \\
\nodata & 0.5323417078167814  & 0.050227726038051694 & 37.19385535484577   & 0.15032160219430055 & -145.05520943675262 & 0.1318726138984278   & 59.6754     & \nodata \\
\nodata & 6.37232292017776    & 0.0137946901051907   & 68.79697898216794   & 0.04264188775089677 & -84.07707033295478  & 0.03680789911821181  & -4.03993    & \nodata \\
\nodata & 0.5021240985953683  & 0.029189592695681964 & -57.9436465188258   & 0.1911215986654146  & 16.919634543207177  & 0.10679177694312522  & -58.6918    & \nodata \\
\nodata & 5.237224518979301   & 0.029752252992649626 & 23.771764671058826  & 0.07292367996213067 & -28.25823161266089  & 0.06344587895614055  & -19.6517    & \nodata \\
\nodata & 3.3089435737776687  & 0.04711084402088657  & -41.45242638019648  & 0.3418156551313586  & -4.510012473794788  & 0.4388646698761767   & -19.7339    & \nodata \\
\nodata & 5.072750731615924   & 0.04188258021899252  & 13.125640321529255  & 0.08925571521229253 & -27.369888362552718 & 0.05289115275902435  & -7.76221    & \nodata \\
\nodata & 0.3185097583003641  & 0.014007935590281349 & 6.6397581995189     & 0.27566940326796724 & -20.42262436999549  & 0.4456232897623583   & -27.3047    & \nodata \\
\nodata & 0.7753680397807614  & 0.01945512547484866  & 7.104977799189683   & 0.16560847226787015 & -5.534895098477748  & 0.11471267191919594  & 15.3108     & \nodata \\
\nodata & 1.4924247624828302  & 0.028700623137541913 & -22.86896549011412  & 0.13153650570900005 & -6.5044933933639575 & 0.0918536954846762   & -1.93386    & \nodata \\
\nodata & 0.3373361076141824  & 0.020908617609888186 & -23.51284581625556  & 0.1281836138293499  & -51.21200709697348  & 0.07993530283388152  & 67.9022     & \nodata \\
\nodata & 0.9508792557786656  & 0.04214796908087648  & -23.92315869160129  & 0.10866864291003138 & -18.19346642794714  & 0.11087251384043836  & -0.941346   & \nodata \\
\nodata & 2.6649412215771706  & 0.02575193557459232  & -11.654902067800947 & 0.0765638732657513  & 55.80843779696036   & 0.07016094187926156  & 15.9426     & \nodata \\
\nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata

\begin{longrotatetable}
\begin{deluxetable*}{cccccccccc}
\tabletypesize{\scriptsize}
\tablecaption{\apogee\ $JHK$ and \gaia\ $G$, $BP$, $RP$ relative photometric magnitudes.}
\tablehead{\colhead{\nodata} & \colhead{$J$} & \colhead{$H$} & \colhead{$K$} & \colhead{$G$} & \colhead{$BP$} & \colhead{$RP$}}   
\startdata
\nodata & 12.83199977874756  & 12.175999641418455 & 11.96500015258789  & 13.415744          & 14.226433           & 12.53435            \\
\nodata & 14.255999565124512 & 13.72599983215332  & 13.40999984741211  & 12.228349          & 12.702468           & 11.614723           \\
\nodata & 11.791999816894533 & 11.104999542236328 & 10.967000007629396 & 10.691218          & 11.139357           & 10.104778           \\
\nodata & 13.706000328063965 & 13.013999938964846 & 12.895999908447266 & 12.420532          & 12.90665            & 11.800449           \\
\nodata & 11.616999626159668 & 10.946000099182127 & 10.751999855041504 & 12.908026          & 13.56541            & 12.142124           \\
\nodata & 11.312999725341795 & 10.663999557495117 & 10.508000373840332 & 10.826382          & 10.85745            & 10.774274           \\
\nodata & 13.128999710083008 & 12.498000144958494 & 12.32699966430664  & 14.534789          & 15.185692           & 13.744579           \\
\nodata & 13.538000106811525 & 12.902999877929688 & 12.762999534606935 & 14.004284          & 14.160915           & 13.288598           \\
\nodata & 12.69499969482422  & 12.036999702453613 & 11.86299991607666  & 14.93699           & 15.395644           & 14.318204           \\
\nodata & 12.53499984741211  & 11.843999862670898 & 11.732000350952148 & 10.015749          & 10.34129            & 9.559692            \\
\nodata & 12.536999702453613 & 11.928000450134276 & 11.682000160217285 & 13.735478          & 14.700515           & 12.765027           \\
\nodata & 11.274999618530273 & 10.717000007629396 & 10.515000343322754 & 11.817871          & 12.467115           & 11.063624           \\
\nodata & 10.21399974822998  & 9.600000381469727  & 9.407999992370604  & 13.178638          & 13.876391           & 12.384415           \\
\nodata & 14.201000213623047 & 13.718000411987305 & 13.46500015258789  & 12.388862          & 12.901998           & 11.731835           \\
\nodata & 13.104000091552734 & 12.43000030517578  & 12.256999969482422 & 8.78063            & 8.920993            & 8.569135            \\
\nodata & 11.628999710083008 & 10.996000289916992 & 10.802000045776367 & 15.246564          & 16.136223           & 14.321473           \\
\nodata & 12.312999725341795 & 11.618000030517578 & 11.38599967956543  & 11.829312          & 12.26109            & 11.258054           \\
\nodata & 12.07800006866455  & 11.48799991607666  & 11.237000465393066 & 13.222831          & 13.763299           & 12.547942           \\
\nodata & 13.682000160217285 & 12.944000244140625 & 12.72700023651123  & 12.266772          & 12.872325           & 11.548589           \\
\nodata & 13.501999855041506 & 12.85099983215332  & 12.64900016784668  & 10.416036          & 10.922305           & 9.781987            \\
\nodata & 13.95300006866455  & 13.29300022125244  & 13.10099983215332  & 13.099073          & 13.74368            & 12.342266           \\
\nodata & 14.246000289916992 & 13.595000267028807 & 13.397000312805176 & 11.854821          & 12.39487            & 11.183031           \\
\nodata & 13.241999626159668 & 12.670999526977539 & 12.432000160217285 & 13.945971          & 14.459273           & 13.295053           \\
\nodata & \nodata & \nodata & \nodata & \nodata & \nodata & \nodata
\enddata
\end{deluxetable*}
\end{longrotatetable}

>>>>

Another open question is about the [Fe/H] validation, and whether we are happy with just the cmd plot of metallicity. I think our ideas for other tests were:
\begin{itemize}
\item Looking at M+M binary pairs: there are 174 pairs and it looks by eye that most of the pair lines are parallel to the main sequence. I don't know if there's some more clever way to see if the pairs lie in the same metallicity space of the cmd, so this is probably not that useful
\item Overplotting some isochrones at varying metallicities--which I'm still working on
\end{itemize}

Also I still have to make the vertical action plot.
}\color{black}

\end{abstract}

@@

\begin{deluxetable*}{rcl}
\tablecaption{Test results table header description.}
\tablehead{\colhead{Column} & \colhead{Unit} & \colhead{Description}}
\startdata
{\tt APOGEE$\_$ID} 	& 	& \apogee\ \zmass\ designation \\
{\tt BP$\_$MAG }		& mag	& \gaia\ $BP$ band photometry	\\
{\tt BP$\_$RP} 		& mag	& \gaia\ $BP-RP$ color	\\
{\tt CHI$\_$MANN} 	& 	& $\chi^2$ fit of Mann-trained \cannon\ model	\\
{\tt CHI$\_$WEST} 	& 	& $\chi^2$ fit of West-trained \cannon\ model	\\
{\tt DEC} 			& deg	& \apogee\ declination angle	\\
{\tt DIST} 			& pc	& Distance (1/$\varpi$)	\\
{\tt FE$\_$H }		& dex	& Mann-trained \cannon\ [Fe/H]	\\
{\tt GAIA$\_$ID} 		&	& \gaia\ identification number \\
{\tt G$\_$ABS} 		& mag	& \gaia\ $G$ band absolute magnitude	\\
{\tt G$\_$MAG} 		& mag	& \gaia\ $G$ band photometry	\\
{\tt H} 				& mag	& \apogee\ $H$ band photometry	\\
{\tt H$\_$ERR} 		& mag	& \apogee\ $H$ band photometry uncertainty	\\
{\tt J} 				& mag	& \apogee\ $J$ band photometry	\\
{\tt J$\_$ERR} 		& mag	& \apogee\ $J$ band photometry uncertainty	\\
{\tt K} 				& mag	& \apogee\ $K$ band photometry	\\
{\tt K$\_$ERR} 		& mag	& \apogee\ $K$ band photometry uncertainty	\\
{\tt PLX} 			& mas	& \gaia\ parallax	\\
{\tt PLX$\_$ERR} 		& mas	& \gaia\ parallax uncertainty	\\
{\tt PMDEC} 			& mas yr$^{-1}$	& \gaia\ declination proper motion	\\
{\tt PMDEC$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ declination proper motion uncertainty	\\
{\tt PMRA} 			& mas yr$^{-1}$	& \gaia\ radial ascension proper motion	\\
{\tt PMRA$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ radial ascension proper motion uncertainty	\\
{\tt RA} 				& deg	& \apogee\ radial ascension angle	\\
{\tt RP$\_$MAG} 		& mag	& \gaia\ $RP$ band photometry	\\
{\tt RV$\_$APOGEE} 	& km s$^{-1}$	& \apogee\ radial velocity	\\
{\tt RV$\_$APOGEE$\_$ERR} 	& km s$^{-1}$	& \apogee\ radial velocity uncertainty	\\
{\tt RV$\_$GAIA} 				& km s$^{-1}$	& \gaia\ radial velocity	\\
{\tt RV$\_$GAIA$\_$ERR}		& km s$^{-1}$	& \gaia\ radial velocity uncertainty	\\
{\tt SPT} 				& 	& West-trained \cannon\ spectral subtype ($M0-M9$)	\\
{\tt TEFF} 				& K	& Mann-trained \cannon\ effective temperature	\\
{\tt Vx} 				& km s$^{-1}$	& Cartesian $x$ velocity in Galactocentric coordnates	\\
{\tt Vy} 				& km s$^{-1}$	& Cartesian $y$ velocity in Galactocentric coordnates	\\
{\tt Vz} 				& km s$^{-1}$	& Cartesian $z$ velocity in Galactocentric coordnates	\\
{\tt X} 				& pc	& Cartesian $x$ position in Galactocentric coordnates	\\
{\tt Y} 				& pc	& Cartesian $y$ position in Galactocentric coordnates	\\
{\tt Z} 				& pc	& Cartesian $z$ position in Galactocentric coordnates	
\enddata
\end{deluxetable*}

%====================================
